Image reconstruction of stellar objects with Intensity Interferometry using Generative Adversarial Networks

### 2025-02-02 Phase_II/draft/Result.tex
Changes by Sarangi

\section{The Reconstructed Image with GAN}
In this section, we will first discuss phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis where multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of $2 \cdot 10^{-4}$, a kernel size of 5x5, and equal noise percentages in original and generated images. The batch size selected is 1, and the discriminator-generator receives equal training.
\subsection{The Predicted Image with Trained GAN}
The success of the GAN in training the model for Intensity Interferometry to reconstruct images of fast-rotating stars is demonstrated in Fig.~\ref{fig:GAN}. The GAN was trained on training datasets for 60,000 steps. After training, the GAN was tested on different validation datasets, producing predicted images of a fast-rotating star. Fig.~\ref{fig:GAN} presents a set of four combined images demonstrating the GAN's performance in reconstructing the shape, size, and brightness distribution of the fast-rotating star using II. The left panel shows the signals covered by six baselines, which serve as the input for the generator to train the GAN. The first middle panel displays the real image, or ground truth, which the discriminator uses to differentiate between real and generated images from the generator. During training, the GAN aims to mimic these ground truth images. The second middle panel depicts the reconstructed image, or predicted image, produced by the trained GAN. This panel illustrates the success of the GAN in image reconstruction using II. The right panel shows the difference between the ground truth and the predicted image. This difference should be minimized for high-precision image reconstruction with the application of GAN on II. The predicted images of Fig.~\ref{fig:GAN} showed positive results, accurately providing visual information about the source's size and shape, as well as the distributed brightness. This image reconstruction was achieved using only six baselines. However, the results can be further improved by increasing the number of telescopes to cover more (u, v) planes, making the existing and upcoming Cherenkov Telescope Array Observatory (CTAO) an ideal candidate in collaboration with existing CTA. 
\subsection{Evaluation of GAN using moments}
The reconstructed images are good enough visually to claim the success of GAN in reconstructing the image using II. However, there is a need for statistical evaluation of this predicted image to guarantee success. So, we use image moments as a statistical method. Image moments are statistical properties that provide information about the reconstructed shape, size, and intensity distribution of the objects in the image. These moments help quantify key features like the position, orientation, and distribution of brightness in the image, allowing for an objective evaluation of how well the predicted image corresponds to the actual object. We will analyze the consistency and accuracy of the GAN-generated image by comparing its moments to those of the ground truth. It provides a reliable framework for evaluating reconstruction quality, as image moments highlight the subtle differences in geometric and intensity properties that might not be visually apparent.

@@

The zeroth order raw moment called the monopole is the total intensity of an image. It sums up all the pixel values across the image, providing an overall intensity value. So, the study of monopole provides the total flux of fast-rotating stars here. According to eqn.~\ref{eqn:Mom}, the monopole of an image is calculated as 
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}
Fig.~\ref{fig:mom1} shows the monopole for 50 different reconstructed images. It shows the linear behavior for the monopole between the ground truth (the real image) along the x-axis and the predicted image (reconstructed image) along the y-axis, which is obvious for different shape-size sources. This result explains the similarity between the total intensity of both images. It ensures that the predicted image has the approximately correct total brightness (the flux) compared to the ground truth. However, monopole does not explain the position, shape, size, and brightness distribution of fast-rotating stars; there is a need for higher-order moments.

The center of mass for the fast-rotating star or any other stellar object is calculated using the centroid (x-centroid and y-centroid). It represents the spatial position of the image and is calculated using first-order raw moment and monopole. The formulation of centroid along the x and y directions is
\begin{equation}
	\begin{aligned}
		&m_x = \frac{\sum_{x} x I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{10}}{M_{00}} \\
		&m_y = \frac{\sum_{y} y I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{01}}{M_{00}}
	\end{aligned}  
\end{equation}
Fig.~\ref{fig:mom2} and Fig.~\ref{fig:mom3} show the comparison of the x-centroid and y-centroid for 50 predicted images with respect to ground truths, respectively. The clustering of centroids in a given scale range for all the results explains that the reconstructed image correctly represents the spatial location of the fast-rotating star in compare with the ground truth.

Further, these calculated centroids will help to study the shape, size, and brightness distribution of fast-rotating stars in terms of higher-order image moments. For that, the central moment of an image is calculated according to
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - m_x)^p (y - m_y)^q I(x, y).
\end{equation}
The sum of $p$ and $q$ defines the order of the central moment.

The second order central moment ($\mu_{11}, \mu_{20}, \mu_{02}$) has been shown in Fig.~\ref{fig:struc}, which is used to study the structure of a fast-rotating star along the line of sight of observation (explain in upcoming subsection). All these three plots explain the linear relation of second-order moments again as for monopole and show the success of the application of GAN to reconstruct the image with II.

The information on brightness distribution is gathered using the skewness of the image by calculating third-order central moment ($\mu_{30}, \mu_{03}, \mu_{21}, \mu_{12}$) of images. Fig.~\ref{fig:moments} shows all third-order moments for the ground truth and reconstructed image. The skewness along the x and y axis ($\mu_{30}, \mu_{03}$) to test the GAN for II are acceptable, which can be seen in Fig.~\ref{fig:mom7} and Fig.~\ref{fig:mom8}, where a linear relation exists between ground truth and predicted image. However, the remaining higher moments $(\mu_{21}, \mu_{12})$ shown in Fig.~\ref{fig:mom9} and Fig.~\ref{fig:mom10} are not in good terms specially $\mu_{12}$. Further improvement is possible and needs to be studied.

\subsection{The reconstructed Parameters for object}
The centroids $(m_x, m_y)$ represent the center of the fast-rotating star and spatial location on the image. However, the calculated second-order central moment defines the orientation, semi-major axis, and eccentricity of the source \citep{teague1980image}. These parameters fully describe the ellipse that fits the image data based on the moments.

The orientation of a fast-rotating star along the line of sight is defined in terms of second-order central moments as
\begin{equation}
	\theta = \frac{1}{2}\arctan \big(\frac{2\mu_{11}}{\mu_{20} - \mu_{02}}\big).
	\label{eqn:orn}
\end{equation}
The semi-major and semi-minor axis of the stellar object will be calculated with the help of second-order central moment again and denoted as a and b here, respectively
\begin{equation}
	\begin{aligned}
		&a = 2\sqrt{mp + \delta} \\
		&b = 2\sqrt{mp - \delta}
	\end{aligned}
	\label{eqn:semi}
\end{equation}
where,
\begin{equation}
	mp = \frac{\mu_{20} + \mu_{02}}{2}
	\label{eqn:mp}
\end{equation}
and
\begin{equation}
	\delta = \frac{\sqrt{4\mu_{11}^2 + (\mu_{20} - \mu_{02})^2}}{2}.	
	\label{eqn:delta}
\end{equation}
With the help of the axis value, the eccentricity of the fast-rotating star is calculated as
\begin{equation}
	e = \sqrt{1 - a/b}.
	\label{eqn:eccen}
\end{equation}
Fig.~\ref{fig:recons} shows all zeroth to third order moments only for one ground and predicted image. In this figure, there is a red curve for both ground truth and predicted image, which is the representation of all second-order moments in terms of Eqn.~\ref{eqn:orn} to Eqn.~\ref{eqn:eccen} as orientation, axis value, and eccentricity. There is a cyan dot point as well, which represents the zeroth order moment for both images. The third-order moment is expressed in terms of arrows in magenta and blue color to view the skewness along the x and y axes.

>>>>

\section{The Reconstructed Image with GAN}
In this section, we will first discuss phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis where multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of $2 \cdot 10^{-4}$, a kernel size of 5x5, and equal noise percentages in original and generated images. The batch size selected is 1, and the discriminator-generator receives equal training.
\subsection{The Predicted Image with Trained GAN}
The success of the GAN in training the model for Intensity Interferometry to reconstruct images of fast-rotating stars is demonstrated in Fig.~\ref{fig:GAN}. The GAN was trained on training datasets for 60,000 steps. After training, the GAN was tested on different validation datasets, producing predicted images of a fast-rotating star. Fig.~\ref{fig:GAN} presents a set of four combined images demonstrating the GAN's performance in reconstructing the shape, size, and brightness distribution of the fast-rotating star using II. The left panel shows the signals covered by six baselines, which serve as the input for the generator to train the GAN. The first, middle panel displays the real image, or ground truth, which the discriminator uses to differentiate between real and generated images from the generator. During training, the GAN aims to mimic this ground truth. The second middle panel depicts the reconstructed image, or predicted image, produced by the trained GAN. This panel illustrates the success of the GAN in image reconstruction using II. The right panel shows the difference between the ground truth and the predicted image. This difference should be minimized for high-precision image reconstruction with application of GAN on II. So, the predicted images of fig.~\ref{fig:GAN} showed positive results, accurately providing visual information about the source's size and shape, as well as the distributed brightness. This image reconstruction was achieved using only six baselines. However, the results can be further improved by increasing the number of telescopes to cover more (u, v) planes.
\subsection{Evaluation of GAN using moments}
After the success of GAN in reconstructing the image using II, there is a need for evaluation of this result. Here, we use the image moments for this work. Image moments are statistical properties that provide information about the shape, size, and intensity distribution of the objects in the image. So, we evaluate images generated by the GAN with the moments to ensure the accuracy and precision of the predicted image.

@@

The zeroth order raw moment, which is the total intensity of an image and is called the monopole. It sums up all the pixel values across the image, providing an overall intensity value. So, the study of monopole provides the total flux of fast-rotating stars here. According to eqn.~\ref{eqn:Mom}, the monopole of a image is calculated as 
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}
The left figure (fig.~\ref{fig:mom1}) in fig.~\ref{fig:cen} shows the monopole for 50 different reconstructed images. It shows the linear behavior for the monopole between the ground truth (the real image) along the x-axis and the predicted image (reconstructed image) along the y-axis, which is obvious for different shape-size sources. This result explains the similarity between the total intensity of both images. It ensures that the predicted image has the approximately correct total brightness (the flux) compared to the ground truth. However, monopole does not explain the shape, size, and brightness distribution of fast-rotating stars. So, we need higher-order moments.

The center of mass for the fast-rotating star or any other stellar object is calculated using the centroid (in x and y directions). These centroids are given in terms of first-order raw moment and monopole as
\begin{equation}
	\begin{aligned}
		&m_x = \frac{\sum_{x} x I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{10}}{M_{00}} \\
		&m_y = \frac{\sum_{y} y I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{01}}{M_{00}}
	\end{aligned}  
\end{equation}
Fig.~\ref{fig:mom2} and fig.~\ref{fig:mom3} show the comparison of centroid along the x and y axis for 50 predicted images with respect to ground truths. The clustering of centroids in a given scale range for all the results explains that the reconstructed image correctly represents the spatial location of the fast-rotating star.

Now, these centroids will help to study the shape, size, and brightness distribution of fast-rotating stars in terms of higher-order image moments. For that, the central moment of an image is calculated according to
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - m_x)^p (y - m_y)^q I(x, y)
\end{equation}

The second order moment ($\mu_{11}, \mu_{20}, \mu_{02}$) has been shown in fig.~\ref{fig:struc}, which is used to study the structure of a fast-rotating star along the line of sight of observation (explain in upcoming subsection). All these three plots explain the linear relation of second-order moments again as for monopole and show the success of the application of GAN to reconstruct the image with II.

The information on brightness distribution can be gathered using the skewness of the image, and for that, there is a need for third order ($\mu_{30}, \mu_{03}, \mu_{21}, \mu_{12}$) central moment of images. Fig.~\ref{fig:moments} shows all third-order moments for the ground truth and reconstructed image. The skewness along the x and y axis ($\mu_{30}, \mu_{03}$) to test the GAN for II are acceptable, which can be seen in fig.~\ref{fig:mom7} and fig.~\ref{fig:mom8} where a linear relation exists between ground truth and predicted image. However, the remaining higher moments $(\mu_{21}, \mu_{12})$ shown in fig.~\ref{fig:mom9} and fig.~\ref{fig:mom10} are not in good terms specially $\mu_{12}$.

\subsection{The reconstructed Parameters for object}
The centroids $(m-x, m_y)$ represents the center of the fast-rotating star. However, the calculated second-order central moment defines the orientation, semi-major axis, and eccentricity of the source \citep{teague1980image}. These parameters fully describe the ellipse that fits the image data based on the moments.

The orientation along the line of sight is defined as
\begin{equation}
	\theta = \frac{1}{2}\arctan \big(\frac{2\mu_{11}}{\mu_{20} - \mu_{02}}\big).
	\label{eqn:orn}
\end{equation}
The semi-major and semi-minor axis will be calculated according to
\begin{equation}
	\begin{aligned}
		&a = 2\sqrt{mp + \delta} \\
		&b = 2\sqrt{mp - \delta}
	\end{aligned}
	\label{eqn:semi}
\end{equation}
where,
\begin{equation}
	mp = \frac{\mu_{20} + \mu_{02}}{2}
	\label{eqn:mp}
\end{equation}
and
\begin{equation}
	\delta = \frac{\sqrt{4\mu_{11}^2 + (\mu_{20} - \mu_{02})^2}}{2}.	
	\label{eqn:delta}
\end{equation}
So, the eccentricity of the fast-rotating star in terms of axis value is 
\begin{equation}
	e = \sqrt{1 - a/b}.
	\label{eqn:eccen}
\end{equation}
Fig.~\ref{fig:recons} shows all zeroth to third order moments only for one ground and predicted image. In this figure, there is a red curve for both ground truth and predicted image which is the representation of all second-order moments in terms of eqn.~\ref{eqn:orn} to eqn.~\ref{eqn:eccen} as orientation, axis value, and eccentricity. There is a cyan dot point as well, which represents the zeroth order moment for both images. There is also a third-order moment, which is expressed in terms of arrows in magenta and blue color to view the skewness along the x and y axes.

### 2025-02-02 Phase_II/draft/conclusion.tex
Changes by Sarangi

\section{Conclusion}
Intensity Interferometry (II), a technique, is emerging after a long gap to overcome the challenges of Amplitude Interferometry on optical wavelength. It is rejuvenated with support from advanced facilities at the Imaging Cherenkov Telescope Arrays (ICTAs). These telescopes, equipped with large aperture areas and photon detectors capable of resolving signals within a nanosecond, are instrumental in capturing high-resolution images \citep{dravins2013optical}. The Major Atmospheric Gamma Imaging Cherenkov Telescope (MAGIC), one of the CTA in the world, is now also being used for II, observing stellar objects with its 17-meter diameter mirror \citep{lorenz2004status}. The telescope has already demonstrated the potential of II by measuring and comparing the diameters of individual stars \citep{abe2024performance}. Other existing arrays (Very Energetic Radiation Imaging Telescope Array System (VERITAS) and High Energy Stereoscopic System (HESS)) are also making progress as II when gamma-ray observation stops \citep{kieda2021veritas, zmija2022optical}. Future observations, made possible by more advanced, sensitive instrumentation and upcoming CTAO, promise to push the boundaries of what we can observe with optical range. However, there is a drawback of II in terms of phase loss of signal, capturing only its magnitude via photon correlation. Here, the challenge of phase retrieval in Intensity Interferometry (II) has been effectively addressed through the application of machine learning techniques, particularly with conditional Generative Adversarial Networks (cGAN). Our study demonstrates that the application of cGAN on II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. The evaluation based on image moments, specifically the monopole, second, and third-order moments, supports the effectiveness of the cGAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines. A critical factor in the performance of the reconstruction process is the extent of Fourier plane coverage, which is determined by the number of available telescopes and the total observing time. The brightness distribution can likely be reconstructed with even higher precision with a full night of observation using four telescopes. Future work could explore different observatory layouts, such as the Southern Cherenkov Telescope Array (CTA) layout, to evaluate the impact on image reconstruction quality.

While the results of this study demonstrate the significant potential of machine learning, particularly cGAN, for image reconstruction in II, several aspects need further refinement. Detector efficiencies, which impact the signal-to-noise ratio (SNR) in real-world scenarios, have yet to be incorporated, and addressing them will be crucial for more accurate SNR estimation. Incorporating these efficiencies in future work will be important for providing a more accurate estimate of the SNR and the robustness of the GAN in real-world applications. While cGAN has proven effective in this study, there are other methods for generating images from image data. Exploring and comparing alternative methods, as well as adapting the GAN architecture itself, could offer improvements. For instance, different loss functions could be implemented and compared to enhance reconstruction quality. Further testing is required to refine the GAN and make it more robust and reliable. However, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

In summary, though cGAN has shown promising results, further testing and improvements are necessary to make the model more robust and reliable in practical applications. Overall, this study highlights the potential of machine learning approaches for advancing phase retrieval and high-resolution image reconstruction in II, suggesting exciting future directions in the application of AI to optical astronomy.

>>>>

\section{Conclusion}
The challenge of phase retrieval in Intensity Interferometry (II) has been effectively addressed through the application of machine learning techniques, particularly with Generative Adversarial Networks (GANs). Our study demonstrates that using a GAN on II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. The evaluation based on image moments, specifically the monopole, second, and third-order moments, supports the effectiveness of the GAN in achieving accurate image reconstruction from a one-night observation using six baselines.

A critical factor in the performance of the reconstruction process is the extent of Fourier plane coverage, which is determined by the number of available telescopes and the total observing time. The brightness distribution can likely be reconstructed with even higher precision with a full night of observation using four telescopes. Future work could explore different observatory layouts, such as the Southern Cherenkov Telescope Array (CTA) layout, to evaluate the impact on image reconstruction quality.

This study has not considered the detector efficiencies, which could influence the signal-to-noise ratio (SNR) in practical scenarios. Incorporating these efficiencies in future work will be important for providing a more accurate estimate of the SNR and the robustness of the GAN in real-world applications.

While conditional GANs have proven effective in this study, there are other methods for generating images from image data. Exploring and comparing alternative methods, as well as adapting the GAN architecture itself, could offer improvements. For instance, different loss functions could be implemented and compared to potentially enhance reconstruction quality. Further testing is required to refine the GAN and make it more robust and reliable. However, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

In summary, while the current implementation shows strong potential, further exploration and refinement of machine learning models are needed to optimize their application in phase retrieval and image reconstruction in Intensity Interferometry.

### 2025-02-02 Phase_II/draft/intro.tex
Changes by Sarangi

Intensity Interferometry (II) emerged in the 1970s as a groundbreaking technique for measuring the diameters, orbits, and limb-darkening coefficients of 32 stars in the N-body system. Despite these significant achievements, the method did not gain widespread adoption due to the then-required highly sensitive photon detectors and sophisticated data analysis techniques. However, advancements in computational methods and modern electronics, which now offer time resolution of detectors in the nanoseconds (ns) range, have renewed II as a feasible technique for resolving astrophysical objects at visible wavelengths. Current simulations with ns resolution of photon detectors show that II is effective in achieving high precision measurement of parameters for stellar objects \citep{10.1093/mnras/stab2391, 10.1093/mnras/stac2433}. Yet, despite these technological advancements, II still faces a significant limitation: the absence of phase information, which is essential for the complete reconstruction of the brightness distribution of stellar sources.

Several theoretical and computational approaches to phase reconstruction with II have been proposed. \cite{gamo1963triple} introduced the concept of triple-intensity correlation, which \cite{goldberger1963use} further applied in an experiment to observe scattered particles in microscopic systems. Sato conducted experiments to measure the diameter and phase of asymmetrical objects, suggesting that triple correlation could extend II to image stellar bodies \citep{sato1978imaging, sato1979computer, sato1981adaptive}. Despite its potential, the sensitivity of the signal to noise poses a significant challenge for this approach.

Later \cite{holmes2010two} proposed an alternative method, utilizing the Cauchy-Riemann relations to reconstruct 1-D images and extending this to 2-D images across a range of signal-to-noise (SNR) values. This algorithm was applied to simulated data of stellar objects using II, considering existing and forthcoming Imaging Cherenkov Telescope Arrays (ICTAs) with a large number of telescopes \citep{nunez2010stellar, nunez2012high, nunez2012imaging}. Despite these theoretical advances, practical application to actual observational data of stellar objects with II has yet to be achieved. There is a need to advance this field by introducing a novel image reconstruction technique.

Currently, artificial intelligence (AI) is applied to high-volume data to reduce human computing resources and time effort. New analysis techniques using Machine Learning (ML) applications on observed data are playing at the front to reveal the astronomical mystery. In radio morphologies, the supervised and unsupervised algorithm is used to distinguish the radio components from the infrared source \citep{wu2019radio, galvin2020cataloguing}. Recently, the image reconstruction of the M87 black hole using Generative Adversarial Networks (GANs) has been performed \citep{10.1093/mnras/stad3797}. There is also a need to introduce the application of ML to II data, which will aim to overcome the limitations of image reconstruction and high-resolution optical astronomy.

In this paper, we propose an alternative approach for image reconstruction of stellar objects with II using GAN \citep{goodfellow2014generative}. We consider four ICTAs and perform a one-night simulation for a fast-rotating star. The predicted image by a trained GAN shows promising results for reconstructing the image's shape and size. Image moments evaluate the predicted image and demonstrate that the size and shape of the fast-rotating star are reconstructed. However, the brightness distribution recovers only up to the third-order moment. With the introduction of more telescopes, we can expect even greater precision, especially in recovering higher-order image moments. This approach holds great promise, pushing the boundaries of high-resolution optical astronomy and opening up new possibilities for studying stellar objects with remarkable detail.

This paper is organized as follows: The first section discusses Intensity Interferometry following its signal and noise for fast-rotating stars along the Earth's rotation. The second section introduces the GAN formulation and its structure. The third section details the parameter selection for training the GAN to reconstruct the image. After that, the fourth section presents the results of trained GAN visually and using image moments. At the end, there is a conclusion of the entire result.

>>>>

Intensity Interferometry (II) \textcolor{blue}{was first reported by Hanbury Brown and Twiss (HBT) during the 1950s \citep{brown1954lxxiv, HBT56} as a \textquotedblleft new type of interferometry\textquotedblright \ to measure stellar parameters such as angular diameter, orbits, limb darkening coefficients etc. Later theoretical results reported by 
\cite{brown1957interferometry, brown1958interferometry}, 
and those of \cite{glauber1963quantum} and 
\cite{MandelWolf1995, Hecht2002} demonstrated the deeper 
physical properties of photon correlations that lay at the 
core of II and also laid the foundation of Quantum Optics.}
By the 1970s, \st{as a groundbreaking} \textcolor{blue}
{with the results of stellar parameter measurements of 32 stars 
in single and multistar systems, 
carried out by Hanbury Brown and his collaborators \citep{hanbury1974angular} at the historic Narrabri Stellar Intensity Interferometer (NSII), Australia,  
II had emerged as an alternative technique 
for measuring stellar parameters.}\st{diameters, orbits, and limb-darkening 
coefficients of 32 stars the N-body system.} Despite these significant achievements,
the method did not gain 
widespread adoption \textcolor{blue}{during several following decades due to unavailability of} the required 
sensitive photon detectors and sophisticated data analysis 
\textcolor{blue}{devices}\st{techniques}. 
However, \textcolor{blue}{subsequent} advancements in computational 
methods and in modern electronics, which currently offer time resolution of detectors in the nanoseconds (ns) to picoseconds (ps) range, have renewed 
II as a feasible technique for resolving astrophysical objects at visible wavelengths. Recent 
 simulations deploying ns-resolution photon detectors show that II 
 \st{is} \textcolor{blue}{could be} effective 
 in achieving high precision measurement of parameters for stellar objects 
 \citep{10.1093/mnras/stab2391, 10.1093/mnras/stac2433}.

In addition to measurement of physical parameters of star systems, a fundamental motivation of optical astronomy is to obtain the sky intensity distribution of the source using the information observed on the ground. In the context of II, the task is to construct the image of the source from the intensity correlations observed by a set of pairs of telescopes (light buckets) on the ground.  \textcolor{blue}{However, with the basic observable of II being the electromagnetic field intensity, and not the field strength},
 a significant aspect of \textcolor{blue}{this method} is the loss of phase information. \st{Consequently, a} A full 
 reconstruction of the brightness distribution of a source with II 
 necessitates phase information. Therefoere the task is to reconstruct the phase of the signal.

Several theoretical and computational approaches to phase reconstruction 
with II have been proposed. \cite{gamo1963triple} introduced the concept 
of triple-intensity correlation, which \cite{goldberger1963use} further 
applied in an experiment to observe scattered particles in microscopic systems. 
Sato conducted experiments to measure the diameter and phase of asymmetrical 
objects, suggesting that triple correlation could extend II to image stellar 
bodies \citep{sato1978imaging, sato1979computer, sato1981adaptive}. \textcolor{blue}{However, achieving a good signal-to-noise ratio (SNR) still poses a significant challenge for this approach.}

Gerchberg and Saxton suggested an iterative method \cite{GerchbergSaxton1972}to determine the phase of from the image and diffraction plane pictures. This method relies on good initial estimates and may struggle with slow convergence issues otherwise. Fienup suggested a Hybrid Input-Output algorithm that introduced feedback mechanisms to improve convergence rate and robustness, particularly in noisy environments.

Later \cite{holmes2010two} proposed an alternative method, utilizing the 
Cauchy-Riemann relations to reconstruct 1-D images and extending this to 
2-D images across a range of signal-to-noise (SNR) values. This algorithm 
was applied to simulated data of stellar objects using II, considering 
existing and forthcoming Imaging Cherenkov Telescope Arrays (IACTs) with 
a large number of telescopes 
\citep{nunez2010stellar, nunez2012high, nunez2012imaging}. This method faces challenge of computational complexity in attempting to generalize to higher dimensions.

Li etal suggested a flexible iterative Regularization method \citep{Li2014} of incorporating prior information (e.g., sparsity, smoothness, or non-negativity), reducing the ill-posedness of the phase retrieval problem. This method is more robust against noise and stabilizes the solution against artefacts and spurious solutions. This method faces challenges of choice of regularization parameter, computational complexity and sensitivity to initial guess.

The Transport-of-Intensity Equation (TIE) method, a non-interferometric technique and first suggested by Teague \citep{Teague1983} relates the intensity variations along the optical axis and phase of the optical fields. It enabled phase retrieval from intensity measurements at multiple planes. Zhang et al suggested a method \citep{Zhang2020} of obtaining a \textquotedblleft universal solution \textquotedblright to TIE using \textquotedblleft maximum intensity assumption \textquotedblleft, hence converting the TIE into a Poisson equation and then by solving it iteratively. Recent work of Kirisits et al \citep{Kirisits2024} have explored hybrid methods combining TIE with other equations, such as the Transport of Phase Equation (TPE). These approaches leverage the strengths of both equations to improve phase retrieval accuracy. This method has universal applicability in the sense that it works for arbitrarily shaped apertures, handles non-uniform illumination and inhomogeneous boundary conditions. This method guarantees convergence although the speed of convergence is dependent on initial guess quality. The final results are dependent on boundary conditions.

\textcolor{red}{With non-linearlity built into its grains, neural networks and deep learning were obvious choice to look for as an option to use. There have been these works.. involving deep learning, GAN.........{\bf Reference and brief description of other pieces of work using NN and Deep Learning to be inserted here!!}}

In this paper, we propose an alternative approach using Generative 
Adversarial Networks (GANs) \citep{goodfellow2014generative} to reconstruct 
images of stellar objects using II. We consider four Imaging Cherenkov 
Telescope Arrays (IACTs) located in the Northern Hemisphere and perform 
a one-night simulation for a fast-rotating star. The predicted image using 
a trained GAN shows promising results for reconstructing the image's shape and 
size.The evaluation of the predicted image is conducted using image moments, 
demonstrating that the size and shape of the fast-rotating star are 
reconstructed. However, the brightness distribution recovers only at the 
third-order moment.

This paper is organized as follows: The first section discusses 
Intensity Interferometry following its signal and noise for fast-rotating 
stars along the Earth's rotation. The second section introduces the GAN 
formulation and its structure. Following the third section details the 
parameter selection for training the GAN to reconstruct the image. 
After that, the fourth section presents the results of trained GAN visually 
and using image moments. At the end, there is a discussion of the entire result.

### 2025-02-02 Phase_II/draft/main.tex
Changes by Sarangi

\documentclass[fleqn,usenatbib,twocolumn]{mnras}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{enumerate}
\usepackage{caption,subcaption}
\usepackage{comment}
\captionsetup{compatibility=false}
\usepackage{animate}
\usepackage{amsmath} % For \text and many other math symbols
\usepackage{amssymb} % For \mathbb
\usepackage{amsfonts} % For \mathbb

@@

\title{Generative AI for Image Reconstruction with Intensity Interferometry: A First Attempt}
\date{}

@@

\begin{abstract}
In modern astronomy, simulation and observation with Intensity Interferometry (II) are showing noteworthy progress with optical wavelengths to achieve high-resolved stellar objects. Despite these achievements, phase retrieval of the signal remains challenging due to photon correlation, hence the image reconstruction. Here, we propose the application of a conditional Generative Adversarial Network (cGAN) to II simulated data to overcome the limitations of traditional image reconstruction methods, as Artificial Intelligence (AI) techniques are at the forefront of revealing astronomical mysteries by analyzing the signal. Here, the cGAN effectively reconstructs the shape, size, and brightness distribution of a fast-rotating star by utilizing sparsely sampled, phase-subtracted Fourier transforms of the star. Simulations of II are based on an assembly of four telescopes similar to existing arrays. However, the sensitivity of the signal and high resolution are expected to improve with additional baselines, which makes the current and future Cherenkov Telescope Array Observatory (CTAO) an ideal candidate for II applications. This approach is highly relevant to the existing image reconstruction technique with II and could revolutionize high-resolution imaging in astronomy.
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}

\bibliographystyle{mnras}
\bibliography{refs.bib}

>>>>

\documentclass[fleqn,usenatbib,twocolumn]{mnras}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{enumerate}
\usepackage{caption,subcaption}
\usepackage{comment}
\captionsetup{compatibility=false}
\usepackage{animate}
\usepackage{amsmath} % For \text and many other math symbols
\usepackage{amssymb} % For \mathbb
\usepackage{amsfonts} % For \mathbb
%%%%%
\usepackage{color}
%\usepackage[normalem]{ulem}
\usepackage{soul}
\usepackage{ulem}
\usepackage{comment}

@@

\title{Image reconstruction of stellar objects with Intensity Interferometry using Generative Adversarial Networks}
\date{}

@@

\begin{abstract}
In current astronomy, the simulation of Intensity Interferometry is demonstrating significant advancements in the high-precision resolution of stellar objects at optical wavelengths. Despite these achievements, phase retrieval remains challenging due to photon correlation. This paper proposes the application of a conditional Generative Adversarial Network (GAN) to address this issue of image reconstruction with Intensity Interferometry. Here, the GAN effectively reconstructs the shape, size, and brightness distribution of a fast-rotating star by utilizing sparsely sampled, phase-subtracted Fourier transforms of the star. This innovative approach promises to enhance the capabilities of Intensity Interferometry for image reconstruction in addition to estimating the parameters with high precision in astronomical observations.
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}

\bibliographystyle{mnras}
\bibliography{refs}
